---
title: "Mídia, identificação partidária e adesão democrática"
author: "Amanda Freitas"
date: '2022-11-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F, warning = F, paged.print = T)
```

```{r include=FALSE}
library(tidyverse)
library(tidylog)
library(tidyr)
library(data.table)
library(readr)
```



Dieta de mídia e uso de mídia para se informar sobre política como VARIÁVEL INDEPENDENTE

Impacto nas variáveis de adesão democrática e partidarismo VARIÁVEL DEPENDENTE

Maior uso de redes sociais e internet está relacionado com menor adesão democrática e antipartidarismo?

A literatura diz que com as redes sociais o consumo de conteúdo político com desinformação e polarizador, incitando ódio e intolerância, aumentou. Isso significa que indivíduos que usam mais esse tipo de mídia do que a mídia tradicional tenderiam a ter opiniões mais extremistas. 

## LATINOBAROMETRO 2020

Q10STGBS. With which of the following statements do you
agree most? (READ THE ALTERNATIVES AND TICK ONE)
Democracy is preferable to any
other kind of government...............1

Under some circumstances, an
authoritarian government can be
preferable to a democratic
one....................................2

For people like me, it doesn’t matter
whether we have a democratic or
non-democratic regime..................3

DNK....................................8

Essa é uma questão clara de ADESÃO DEMOCRÁTICA que pode ser modelada como variável discreta -> o que leva as pessoas a escolherem?

Q11STGBS.A-B In general, would you say you are very
satisfied, quite satisfied, not very satisfied, or not
at all satisfied with the working of the democracy in
(country)? (READ ALTERNATIVES AND TICK ONE IN
Q11STGBS.A)  
Q11STGBS.
Very satisfied ........ 1
Quite satisfied ....... 2
Not very satisfied .... 3
Not at all satisfied .. 4

Isso aqui importa pra adesão democrática, uma pessoa mais satisfeita tende a dizer que está mais feliz com a democracia?



Q18ST. (SHOW CARD 4) In politics, people normally speak of “left” and “right”. On a scale where 0 is left and 10 is right, where would you place yourself? (TICK ONE ANSWER)
LEFT RIGHT
00
01
02
03
04
05
06
07
08
09
10
None...97 DNK...98 NA...99DON’T READ

Ser de esquerda ou de direita importa? eu posso transformar em variável discreta com 3 categorias - esquerda, centro e direita

Q20STA-20STG. (SHOW CARD 6) Do you strongly agree (1), agree (2), disagree (3) or strongly disagree (4) with the following statements. (READ EACH STATEMENTS AND TICK ONE FOR EACH)
SA A D SD DNK/NA
Q20ST.A Democracy may have
problems but it is the best
system of government. .................. 1 2 3 4 0
Q20STM.B I wouldn't mind if an
undemocratic government came to power
if it can solved the problems. ........ 1 2 3 4 0
Q20STM.C Democracy allows us to
solve the problems we have ............. 1 2 3 4 0
Q20STM.D In case of difficulties it is
allright for the president to control
the media .............................. 1 2 3 4 0
Q20ST.E It is possible to eradicate
corruption from politics ............... 1 2 3 4 0
Q20ST.F. It is allright to pay the
price of some corruption
if problems are solved. ............... 1 2 3 4 0
Q20STI.G. If I do not report an act of
corruption that I am aware of,
I become an accomplice ................. 1 2 3 4 0

Mais uma variável de adesão democrática

Q46ST. How interested are you in politics? (READ ALTERNATIVES AND TICK ONE)
Very interested............ 1
Fairly interested.......... 2
A little interested.........3
Not at all interested ..... 4
DNK................. 8 DON’T READ
DNA............... 0

INTERESSE EM POLÍTICA, pode ser transformado tb em uma dummy - interessado ou nao interessado - ser interessado faz a pessoa aderir mais à democracia??



## VARIÁVEIS MIDIÁTICAS
Eu posso fazer com vars categóricas, traçando por exemplo 4 perfis de consumo de mídia 




